\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-18 pldi/draft.tex
Finish rewrite

At a high level, \sys takes as input a POSIX shell script like the one in Fig.~\ref{fig:example} and outputs another shell script.
The new script may incorporate parallelism, and handed off to the user's original shell interpreter---\eg \ttt{bash} or \ttt{zsh}---for execution.
To discover and enable parallel execution, \sys analyzes individual program fragments, addressing the aforementioned challenges.

The key insight behind the analysis is that, intuitively, the POSIX language amounts to a few constructs that indicate different levels of candidacy for pararllelization.
For example,
 (i) the parallel-composition operator exposes task parallelism,
 (ii) commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
 (iii) the pipe operator is a candidate for both task and data parallelism, and
 (iv)  other composition operators can be viewed as barrier candidates.

Second, \sys needs to understand the parallelization characteristics of individual commands used to compose larger programs~\sx{parallelizability}.
This challenge can be broken down into two parts.
First, understand standard commands part of the shell---that is, built-ins that come with any shell and support its core functions.
One example is the set of commands defined by the \textsc{POSIX} standard~\cite{posix}.
\sys addresses this part by a careful study of a group of commands, identifying a small but clear set of well-understood parallelizability classes~\sx{cmd}.
This group is already useful-enough to allow the composition of practical pipelines~\sx{micro}.

will have several fragments that explicitely execute in parallel, as long as \sys was able to identify and extract parallelism latent in the original script.

By parsing the script of Fig.~\ref{fig:example},

\sys will first identify a series of commands

For now, different \ttt{for} branches can proceed in parallel as all the information is available.
However, if the 
In cases where \sys does not have enough information in place, it will not attempt parallelizing the script.

\kk{Here we should reiterate on the fact that \sys takes a script and outputs a script.}

\kk{Make sure for the running script to contain more than just a
  pipeline, so that it is clear that \sys turns a script into the same
  script with DFGs in place of ``parallelizable-regions'' (which are
  usually pipelines).}

As outlined earlier, the key insight behind \sys is that the shell already describes parallel computations over streams.

% However, understanding the shell's primitives is not enough
The second part is understanding the characteristics of commands that fall outside the built-in set, allowing new commands to leverage \sys's power.
New commands are arbitrary programs, but still need to conform to the shell's pipeline interface.
To solve this part, \sys leverages the aforementioned study~\sx{cmd} to define a small but expressive domain-specific language (DSL) for describing a command's key properties with respect to its parallelizability.
This DSL is aimed towards developers of commands (not users of the shell), who can quickly and easily capture the parallelizability of the commands they are developing.
% We haven't introduced the notion of flags
% A command developer can only annotate the default behavior;
%   if a new flag is added a few years later, and if that flag alters the command's class, only then a new annotation is needed.

Given a pipeline of commands and their parallelizability characteristics, \sys's next challenge is to analyze it and identify candidate subexpressions for distribution~\sx{ir}.
It addresses this challenge with a series of rewriting passes that perform dataflow analysis to identify distributable regions.
It converts the AST to a dataflow graph, and iteratively performs graph transformations that expose data parallelism as task parallelism, while preserving the pipeline's correctness.
These transformations are complicated by the shell's pervasive use of dynamic features---\eg environment variables, file input sizes, \etc

To schedule and execute the resulting distributed dataflow graph, \sys needs to solve several challenges~\sx{impl}.
First, it needs to wrap single-node commands with corrective runtime interposition based on their classes.
This is achieved by thin wrappers that wrap all the commands involved in the program, generated using the parallelizability specification described earlier.
Thin wrappers avoid interference with legacy functionality, as \unix commands contain non-obvious logic and flags added over extended periods of time.

Second, \sys needs to fill-in planing details left blank by the rewriting pass.
This is achieved by \emph{planner activators}, shell expressions that yield (and pass their arguments) back to the planner.
They are inserted by the planner and evaluated purely for their side-effects (\ie yielding to the scheduler) rather than their return value (which is empty, not affecting pipelines).
They make clever use of shell features such as expansion to achieve the latest possible binding, when critical runtime information becomes available.
% From the point of view of the shell, planner activators are expressed through a higher-order command---similar to \ttt{xargs} and \ttt{time}---that takes other commands as arguments.

% FIXME \kk{I am not exactly sure I understand that :) } \nv{which part?}

The next few sections (\S\ref{parallelizability}--\ref{impl}) discuss \sys's design and implementation details.
% They also outline other, less obvious challenges---such as the synchronization of environment variables, distributed file operation, \etc.

@@

\TODO{Change naming to parallelizability}

>>>>

At a high level, \sys takes as input a POSIX shell script like the one in Fig.~\ref{fig:example} and outputs a new POSIX shell script.
The new script may incorporate parallelism, and is handed off to the user's original shell interpreter---\eg \ttt{bash} or \ttt{zsh}---for execution.
To discover and enable parallel execution, \sys analyzes individual program fragments, addressing the challenges mentioned earlier.

\sys first needs to understand the POSIX constructs used to compose larger programs, indicating indicate different levels of candidacy for pararllelization.
To achieve this, \sys breaks POSIX constructs into a few broad classes:
 (i) the parallel-composition operator (\ttt{\&}) exposes task parallelism,
 (ii) special commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
 (iii) the pipe operator (\ttt{|}) is a candidate for both task and data parallelism, and
 (iv)  other composition operators such as \ttt{\&\&}, \ttt{||}, and \ttt{;} are barrier candidates.

Second, \sys needs to understand the parallelization characteristics of individual commands used to compose larger programs---a two-fold challenge.
To understand standard commands available in any shell, \sys groups POSIX and GNU commands into a small but clear set of well-understood parallelizability classes;
  these commands are already useful-enough to allow the composition of useful shell programs operating in practice.
To understanding the characteristics of commands that fall outside the built-in set, allowing new commands to leverage \sys's power, \sys leverages this study to define an annotation language for describing a command's key properties with respect to its parallelizability.
Such light annotations are aimed towards command developers (rather than users), who can quickly and easily capture the properties of the commands they are developing.
% We haven't introduced the notion of flags
% A command developer can only annotate the default behavior;
%   if a new flag is added a few years later, and if that flag alters the command's class, only then a new annotation is needed.

% Given a pipeline of commands and their parallelizability characteristics, \sys's next challenge is 
% \sys achieves this through a series of rewriting passes that perform dataflow analysis to identify distributable regions.
% It converts the AST to a dataflow graph, and iteratively performs graph transformations that expose data parallelism as task parallelism, while preserving the pipeline's correctness.

Provided information about the shell constructs and individual commands and their flags, \sys needs to analyze the given script and identify program fragments that are candidates for parallelization.
To achieve this, it converts the script into a data-flow graph (DFG) model.
% Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
This is a flexible representation that allows a series of parallelization transformations to convert the graph into its parallel equivalent, in which DFGs indicate the parallelizable regions.
Further transformations compile the DFG back to a parallel script that uses POSIX shell constructs to guide parallelism explicitly---while aiming at preserving the semantics of the sequential program.

Finally, \sys needs to provide correct, high-performance combinators for certain classes commands, and be proactive about (addressing) several practical challenges related to quirks of the \unix environment.
In terms of combiners, \sys offers a set of linkable components for stream splitting and merging;
  these commands live in the \ttt{PATH} and are addressable by name, which means they can be used like (and by) any other commands.
Practical challenges---related to task parallelism, deadlock prevention, and runtime performance---are addressed through a series of constructs that \sys provides.

Returning to the script of Fig.~\ref{fig:example}...\nv{I think we should show fragments of the resulting script}

The next few sections (\S\ref{parallelizability}--\ref{impl}) discuss the details.
% They also outline other, less obvious challenges---such as the synchronization of environment variables, distributed file operation, \etc.

@@
